---
title: "Real Estate Analysis" 
author: "Navya Annapareddy"
fontsize: 12pt
geometry: margin=1in
urlcolor: black
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, highlight=FALSE)
#library(tidyverse)
library(ggplot2)
library(ggpubr)
require(dplyr) 
library(ade4)
library(lattice) 
library("GGally")
library(MASS)
library(lindia)
```

```{r}
house = read.csv("~/Desktop/Fall 2020/SYS4021/Source/housing.csv",sep=',',header=T)
head(house)
```

```{r}
boxplot(house$price, plot=FALSE)$out

```

```{r}
boxplot(house$price, plot=FALSE)$stats[5]
```

```{r}
ggplot(df, aes(x="", y=price)) + 
  geom_boxplot(width=0.6) +
  stat_summary(
    aes(label=sprintf("%1.1f", ..y..)),
    geom="text", 
    fun.y = function(y) boxplot.stats(y)$stats,
    position=position_nudge(x=0.33), 
    size=3.5) +
  theme_bw()
```

```{r include=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- round(abs(cor(x, y, use = "complete.obs")), 2)
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = 2)
}

panel.hist <- function(x, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y,
    col="steelblue2", ...)
}


uva.pairs <- function(vars, ...)
{
	args <- list(...)
	
	if(is.matrix(vars) | is.data.frame(vars)){
		if(is.null(args$labels))pairs(vars, lower.panel = panel.smooth, upper.panel = panel.cor, diag.panel = panel.hist, main = args$main)
		else(pairs(vars, lower.panel = panel.smooth, upper.panel = panel.cor, diag.panel = panel.hist, main = args$main, labels = args$labels))
		}
	else(if(is.character(vars)){
		if(is.null(args$labels))pairs(formula(vars), lower.panel = panel.smooth, upper.panel = panel.cor, diag.panel = panel.hist, main = args$main, data = args$data)
		else(pairs(formula(vars), lower.panel = panel.smooth, upper.panel = panel.cor, diag.panel = panel.hist, main = args$main, data = args$data, labels = args$labels))} 
	else(cat("You must enter a matrix, dataframe or formula")))
	}
```

```{r}
uva.pairs(house[,c("price", "sqft", "bedrooms", "baths")])
```

```{r}
ggplot(house, aes(x=City, y=price)) + geom_point()
```

```{r}
house.main <-lm(price~sqft+bedrooms+baths,data=house)
summary(house.main)
```

0.0000000006526 < 0.01

For every one unit increase in  the predicted value of  increases by the value of the slope. 

```{r}
house.inter <-lm(price~(sqft + bedrooms + baths)^2,data=house)
summary(house.inter)
```

```{r}
anova(house.main,house.inter)
```

p > 0.05
Do not significantly effect 
Therefore reduce complexity and choose smaller

```{r}
anova(house.inter, house.main)
```

```{r}
AIC(house.main)
#Adjusted r^2 = 0.416
AIC(house.inter)
#Adjusted r^2 = 0.4089
```

Choose smaller model based on
- F test p > 0.05
- Smaller AIC
- Larger R^2

Choose larger model based on
- None

```{r}
gg_cooksd(house.main, label = TRUE, show.threshold = TRUE,
          threshold = "convention") +
  xlab("Obs. Number")+ylab("Cook's distance") +
  ggtitle("Cook's distance plot")+
  theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5))
```

```{r}
plot(house.main,labels.id = NULL, which=4) 
```

```{r}
# Residual vs. Fitted
plot(house.main,which=1) 
```
No relationship above/below line = good fit
Constant variance

```{r}
plot(house.main,which=2) 
```

```{r}
boxcox(house.main) 
```

#You should see if the confidence interval includes 1 or 0. If it includes 1, you don't need a transformation. If it includes 0, you should use a log transformation. If it includes neither, use the optimal lambda. 

```{r}
boxcox(house.main, plotit = F)$x[which.max(boxcox(house.main, plotit = F)$y)] 
```

```{r}
City_fact <- as.factor(df$City)
contrasts(City_fact)
```

```{r}
length(unique(df$City))
```

```{r}
pca_df <- df[,c("price", "sqft", "bedrooms", "baths")]
pca_df.corr <- princomp(pca_df, cor=T)

barplot(pca_df.corr$loadings[,1], main='PC1 Loadings with Correlation Matrix')
```

```{r}
names(df)
```

```{r}
cumplot <- function(pca.obj, ...)
{
  xc <- cumsum(pca.obj$sdev^2)/sum(pca.obj$sdev^2)
  barplot(xc, ylim = c(0,1), main = "Proportion of Variance", ylab = "Proportion", names.arg = 1:length(pca.obj$sdev), xlab = "Components", ...)
  xc <- as.data.frame(xc)
  setDT(xc, keep.rownames=TRUE)[]
  names(xc)[names(xc) == "rn"] <- "Component"
  names(xc)[names(xc) == "xc"] <- "Proportion"
  return(xc)
}

cumplot(pca_df.corr, col = "grey")
```

```{r}
biplot(pca_df.corr, main="Biplot")
```
